Don't Mind the Gap: Bridging Network-wide Objectives and Device-level Configurations

### 2016-06-21 paper/v2/paper-edited.tex
correcting spelling mistakes

To avoid this problem, the operator may disallow "valley" paths, i.e., those that go up, down, and back up again. This guard can be implemented by $X$ and $Y$ rejecting paths through the other. But that creates a different problem in the face of failures---an aggregation-induced blackhole~\cite{route-aggregation}. If links D--A and X--C fail, X will hear an announcement for \CD{PG2} from D and will thus announce \CD{PG} externally. This announcement will bring to X traffic for \CD{PG1} as well, but because valleys are disallowed, X does not have a valid route for \CD{PG1} and will drop all traffic for it despite the fact that a valid path exists through $Y$.

Thus, we see that devising a configuration that ensures policy compliance in the face of failures is complex and error-prone. \sysname lets operators implement their high-level policy specification in a way that guarantees compliance under all failures if possible---otherwise, it generates a compile-time error. For aggregation, it also provides a lower bound to operators on the number of failures under which aggregation will not result in blackholes.

@@

%\para{Syntax}
The \sysname FE is just a thin layer atop the RIR for describing preference-based path constraints.
Figure~\ref{fig:rir-syntax} shows the RIR syntax. A policy has one or more constraints, each of which has a test on the type of route and a corresponding set of preferred regular paths. Regular paths are regular expressions where the base characters are AS numbers representing either a router or an external AS. Special \In{} and \Out{} symbols refer to any internal or external location respectively. In addition, $\Sigma$ refers to any symbol. We also use the standard regular expression abbreviation $r^+$ for $r \cdot r^*$, a sequence of one or more occurrences of $r$. Predicates ($t$) consist of logical boolean connectives (and, or, not) as well as tests that match a particular prefix (or group of prefixes) and tests for route advertisments with a particular community value (i.e., an integer value associated with a path).
%% For example, the constraint
%% \[(\textit{prefix}=74.3.28.0/24) \Link \\
%% (as200 \cdot \In^+) \Prefer (as100 \cdot \In^+)
%% \]
%% describes a more-preferred set of paths for traffic announced by a prefix no less specific than \CD{74.125.28.0/24}, which starts at AS 200, before entering and staying inside the user's network to get to the destination, and a less-preferred set of paths that start at AS 100 and are otherwise the same. The plus operator $\In^+$ stands for $\In \cdot \In^*$, at least one internal hop. Tests over route types use standard boolean connectives, and can refer to both prefixes and route community values.

@@

The next step rewrites the high-level constraints such as $\Enter$ according to the equivalences in Figure~\ref{fig:rir-syntax}. Since preferences can only occur at the outermost level for an RIR expression, the final step is to ``lift" occurrences of the preference operator in each regular expression. For example, the regular expression $a \cdot (b \Prefer c) \cdot d$ is lifted to $(a \cdot b \cdot d) \Prefer (a \cdot c \cdot d)$ by distributing the preference over the sequence operator. In general, we employ the follwing distributivity equivalences:
%
\[
\begin{array}{c}
  x \odot (y_1 \Prefer \dots \Prefer y_n) = (x \odot y_1) \Prefer \dots \Prefer (x \odot y_n) \\
  (y_1 \Prefer \dots \Prefer y_n) \odot x = (y_1 \odot x) \Prefer \dots \Prefer (y_n \odot x)
\end{array}
\]
%
where $\odot$ stands for an arbitrary regular binary operator, and $x$ is a policy with a single preference. The compiler flags preferences nested under a unary operator (i.e., \textit{star} or \textit{negation}) as invalid.

@@

Intuitively, a PGIR tracks the policy states of each automaton as route announcements move between locations.
%
Consider the topology in Figure~\ref{fig:example-compilation}. Suppose we want a primary route from neighbor W that allows traffic to enter the network at $A$ and utilize the C--D link before leaving the network (through $X$ or $Y$). As a backup, we also want to allow traffic to enter the network from $B$, in which case the traffic can also utilize the C--E link before leaving the network. For simplicity, we assume that the route ends in either $X$, $Y$, or $Z$. The RIR for the policy could be written as:
%
$$(\CD{W} \cdot \CD{A} \cdot \CD{C} \cdot \CD{D} \cdot \Out) \Prefer (\CD{W} \cdot \CD{B} \cdot \In^+ \cdot \Out)$$
%
Figure~\ref{fig:example-compilation} shows the policy automata for each regular expression preference. Since we are interested in the flow of control messages, the automata match backwards.
%
The figure also shows the PGIR after intersecting the topology and policy automata. Every path in the PGIR corresponds to a concrete path in the topology. In particular, every path through the PGIR that ends at a node $v$ such that the preference function $P(v) = \set{i_1, \dots, i_m}$ is non-empty, is a valid topological path that satisfies the policy constraints and results in a particular path with preferences $i_1$ through $i_m$.
%
For example, the path $X \cdot D \cdot C \cdot A \cdot W$ is a valid path in the topology that BGP route announcements might take, which would lead to obtaining a path with the lowest (best) rank of $1$.
BGP control messages can start from peer X, which would match the $\Out$ transition from both automata, leading to state $1$ in the first automaton, and state $1$ in the second automaton. This possiblity is reflected in the product graph by the node with state $(X,1,1)$. From here, if X were to advertise this route to D, it would result in the path $D \cdot X$, which would lead to state $2$ in the first automaton, and state $2$ in the second automaton, and so on.
%
The ``--" state indicates the corresponding automaton cannot accept the current path or any extension of it. Since
node $(W,5,-)$ is in an accepting state for the first automaton, it indicates that this path has preference 1.

\para{Minimization}
After building the PGIR as desribed above,
we minimize it in order to improve the subsequent analysis that checks if the policies captured by it are safe under failures.
The minimization is based on the observation that, although every path in the PGIR is a valid path in the topology, we do not want to consider paths that form loops. In particular, BGP's loop prevention mechanism forces an AS to reject any route that is already in the AS path.
%
For example, in Figure~\ref{fig:example-compilation}, the path $W \cdot A \cdot C \cdot B \cdot W$ is a valid topological path, leading to a path that satisfies the preference 1 policy, but which contains a loop.
%
The precision of the failure-safety analysis often increases when nodes and edges that are not possible due to loops are eliminated.
%For example, if we can safely remove a PGIR edge, then we do not have to consider what happens when the edge has failed.

We use graph dominators~\cite{tarjan-dominance} as a cheap approximation for removing many nodes and edges in the PGIR that are never on any \emph{simple} (loop free) path between the start and end nodes.
In the PGIR, a node $x$ dominates a node $y$ if $x$ appears on every path leading from the start node to $y$. Similarly, a node $x$ post-dominates a node $y$ in the PGIR if $x$ appears on every path from $y$ to the end node.
We can safely remove nodes and edges in the PGIR when any of the following conditions hold, where $x$, $x'$ and $y$, $y'$ refer to PGIR nodes that are shadows of the same topology location.
\begin{itemize}
\setlength{\itemsep}{1pt}
\setlength{\parskip}{0pt}
\setlength{\parsep}{0pt}
\item Remove $x$ if it is not reachable from the start node
\item Remove $x$ if it can not reach the end node
\item Remove $x$ if it is (post-)dominated by some $x'$
\item Remove edge ($x$, $y$) if $y$ is post-dominated by some $x'$
\item Remove edge ($x$, $y$) if $x$ is dominated by some $y'$
\end{itemize}
For example, node $(W,1,1)$ in Figure~\ref{fig:example-compilation} is removed since it is never on a simple path to the end node, because the path must always go through node $(W,-,4)$. That is, node $(W,-,4)$ post-dominates node $(W,1,1)$ and both are shadows of topology location $W$.
Similarly, the edge from $(C,3,2)$ to $(D,-,2)$ is removed since node $(C,-,2)$ post-dominates $(D,-,2)$ and is a shadow of the same topology location $C$.
We repeatedly apply the minimizations above until no further minimization is possible.
In the example from Figure~\ref{fig:example-compilation}, colored nodes and dashed edges show edges and nodes removed after minimization.

@@

For example, recall the datacenter example from \S\ref{sec:motivation}, with the policy
\CD{PG1} \Path \text{ }\End(\CD{A}), where \CD{PG1} falls under the \CD{PG} aggregate. Figure~\ref{fig:aggregation-safety} shows the simplified PGIR. Since the compiler knows aggregation will occur at $X$, and it knows that the \CD{PG1} prefix will originate at $A$, we can compute the number of failures it would take to disconnect $A$ from $X$. We could remove the $A$--$D$--$X$ path first. We would then need to remove any other $A$--$D$ or $D$--$X$ links from the PGIR (in this case none). Next, we could remove the links along the $A$--$C$--$X$ path, repeating the process. Because $A$ is then disconnected from $X$, the compiler knowns that 2 is a lower bound on the number of failures required to potentially introduce an aggreagation black hole for prefix \CD{PG1}. This process is repeated for other aggregation locations (e.g., $Y$).

@@

Since the compiler can control community tagging only for routers under the control of the AS being programmed, it cannot match on communities for external ASes. Instead, it translates matches from external ASes into a BGP regular expression filter. For example, node $D$ in Figure~\ref{fig:example-compilation} would match the single hop external paths $X$ or $Y$. In general, if routes are allowed from beyond $X$ or $Y$, these will also be captured in the BGP regular expression filters. The unknown AS topology is modelled as a special node in the PGIR that generates a filter to match any sequence of ASes.

@@

After configuration generation, the compiler further processes the ABGP policy, removing community tags when possible, combining filters, removing dead filters, and so on. In the compilation example all community tags can be removed, since there is never any ambiguity based on the router importing the route and the neigbor the route is being imported from. Similarly, after removing the communities, the two rules at router $B$ are merged into a single rule.

@@

Our \sysname compiler is implemented in roughly 6700 lines of F\# code. It includes command-line flags for enabling or disabling the use of the BGP MED attribute, AS path prepending, the no-export community, as well as for ensuring at least k-failure safety for aggregate prefixes. Since each prefix has a separate routing policy, we compile each routing policy in parallel. Currently, \sysname supports generating Quagga router configurations out of the box. Users can add new vendor-specific adaptors to translate from ABGP to other router configuration languages, or incorporate the compiler into an existing template-based system, e.g., by mixing the \sysname-generated BGP configuration with other, non-BGP configuration elements.

@@

\section*{Acknowledgements}

\ryan{TODO: add acknowledgements}
We thank R. Aditya, George Chen, and Lihua Yuan for feedback on this work and the SIGCOMM reviewers for comments on the paper.
This work is supported in part by
the NSF award CNS-1161595.

>>>>

To avoid this problem, the operator may disallow "valley" paths, i.e., those that go up, down, and back up again. This guard can be implemented by $X$ and $Y$ rejecting paths through the other. But that creates a different problem in the face of failures---an aggregation-induced black hole~\cite{route-aggregation}. If links D--A and X--C fail, X will hear an announcement for \CD{PG2} from D and will thus announce \CD{PG} externally. This announcement will bring to X traffic for \CD{PG1} as well, but because valleys are disallowed, X does not have a valid route for \CD{PG1} and will drop all traffic for it despite the fact that a valid path exists through $Y$.

Thus, we see that devising a configuration that ensures policy compliance in the face of failures is complex and error-prone. \sysname lets operators implement their high-level policy specification in a way that guarantees compliance under all failures if possible---otherwise, it generates a compile-time error. For aggregation, it also provides a lower bound to operators on the number of failures under which aggregation will not result in black holes.

@@

%\para{Syntax}
The \sysname FE is just a thin layer atop the RIR for describing preference-based path constraints.
Figure~\ref{fig:rir-syntax} shows the RIR syntax. A policy has one or more constraints, each of which has a test on the type of route and a corresponding set of preferred regular paths. Regular paths are regular expressions where the base characters are AS numbers representing either a router or an external AS. Special \In{} and \Out{} symbols refer to any internal or external location respectively. In addition, $\Sigma$ refers to any symbol. We also use the standard regular expression abbreviation $r^+$ for $r \cdot r^*$, a sequence of one or more occurrences of $r$. Predicates ($t$) consist of logical boolean connectives (and, or, not) as well as tests that match a particular prefix (or group of prefixes) and tests for route advertisements with a particular community value (i.e., an integer value associated with a path).
%% For example, the constraint
%% \[(\textit{prefix}=74.3.28.0/24) \Link \\
%% (as200 \cdot \In^+) \Prefer (as100 \cdot \In^+)
%% \]
%% describes a more-preferred set of paths for traffic announced by a prefix no less specific than \CD{74.125.28.0/24}, which starts at AS 200, before entering and staying inside the user's network to get to the destination, and a less-preferred set of paths that start at AS 100 and are otherwise the same. The plus operator $\In^+$ stands for $\In \cdot \In^*$, at least one internal hop. Tests over route types use standard boolean connectives, and can refer to both prefixes and route community values.

@@

The next step rewrites the high-level constraints such as $\Enter$ according to the equivalences in Figure~\ref{fig:rir-syntax}. Since preferences can only occur at the outermost level for an RIR expression, the final step is to ``lift" occurrences of the preference operator in each regular expression. For example, the regular expression $a \cdot (b \Prefer c) \cdot d$ is lifted to $(a \cdot b \cdot d) \Prefer (a \cdot c \cdot d)$ by distributing the preference over the sequence operator. In general, we employ the following distributivity equivalences:
%
\[
\begin{array}{c}
  x \odot (y_1 \Prefer \dots \Prefer y_n) = (x \odot y_1) \Prefer \dots \Prefer (x \odot y_n) \\
  (y_1 \Prefer \dots \Prefer y_n) \odot x = (y_1 \odot x) \Prefer \dots \Prefer (y_n \odot x)
\end{array}
\]
%
where $\odot$ stands for an arbitrary regular binary operator, and $x$ is a policy with a single preference. The compiler flags preferences nested under a unary operator (i.e., \textit{star} or \textit{negation}) as invalid.

@@

Intuitively, a PGIR tracks the policy states of each automaton as route announcements move between locations.
%
Consider the topology in Figure~\ref{fig:example-compilation}. Suppose we want a primary route from neighbor W that allows traffic to enter the network at $A$ and utilize the C--D link before leaving the network (through $X$ or $Y$). As a backup, we also want to allow traffic to enter the network from $B$, in which case the traffic can also utilize the C--E link before leaving the network. For simplicity, we assume that the route ends in either $X$, $Y$, or $Z$. The RIR for the policy could be written as:
%
$$(\CD{W} \cdot \CD{A} \cdot \CD{C} \cdot \CD{D} \cdot \Out) \Prefer (\CD{W} \cdot \CD{B} \cdot \In^+ \cdot \Out)$$
%
Figure~\ref{fig:example-compilation} shows the policy automata for each regular expression preference. Since we are interested in the flow of control messages, the automata match backwards.
%
The figure also shows the PGIR after intersecting the topology and policy automata. Every path in the PGIR corresponds to a concrete path in the topology. In particular, every path through the PGIR that ends at a node $v$ such that the preference function $P(v) = \set{i_1, \dots, i_m}$ is non-empty, is a valid topological path that satisfies the policy constraints and results in a particular path with preferences $i_1$ through $i_m$.
%
For example, the path $X \cdot D \cdot C \cdot A \cdot W$ is a valid path in the topology that BGP route announcements might take, which would lead to obtaining a path with the lowest (best) rank of $1$.
BGP control messages can start from peer X, which would match the $\Out$ transition from both automata, leading to state $1$ in the first automaton, and state $1$ in the second automaton. This possibility is reflected in the product graph by the node with state $(X,1,1)$. From here, if X were to advertise this route to D, it would result in the path $D \cdot X$, which would lead to state $2$ in the first automaton, and state $2$ in the second automaton, and so on.
%
The ``--" state indicates the corresponding automaton cannot accept the current path or any extension of it. Since
node $(W,5,-)$ is in an accepting state for the first automaton, it indicates that this path has preference 1.

\para{Minimization}
After building the PGIR as described above,
we minimize it in order to improve the precision of the subsequent analysis that checks if the policies captured by it are safe under failures.
The minimization is based on the observation that, although every path in the PGIR is a valid path in the topology, we do not want to consider paths that form loops. In particular, BGP's loop prevention mechanism forces an AS to reject any route that is already in the AS path.
%
For example, in Figure~\ref{fig:example-compilation}, the path $W \cdot A \cdot C \cdot B \cdot W$ is a valid topological path, leading to a path that satisfies the preference 1 policy, but which contains a loop.
%
%The precision of the failure-safety analysis often increases when nodes and edges that are not possible due to loops are eliminated.
%For example, if we can safely remove a PGIR edge, then we do not have to consider what happens when the edge has failed.

We use graph dominators~\cite{tarjan-dominance} as a cheap approximation for removing many nodes and edges in the PGIR that are never on any \emph{simple} (loop free) path between the start and end nodes.
In the PGIR, a node $x$ dominates a node $y$ if $x$ appears on every path leading from the start node to $y$. Similarly, a node $x$ post-dominates a node $y$ in the PGIR if $x$ appears on every path from $y$ to the end node.
We can safely remove nodes and edges in the PGIR when any of the following conditions hold, where $x$, $x'$ and $y$, $y'$ refer to PGIR nodes that are shadows of the same topology location.
\begin{itemize}
\setlength{\itemsep}{1pt}
\setlength{\parskip}{0pt}
\setlength{\parsep}{0pt}
\item Remove $x$ if it is not reachable from the start node
\item Remove $x$ if it can not reach the end node
\item Remove $x$ if it is (post-)dominated by some $x'$
\item Remove edge ($x$, $y$) if $y$ is post-dominated by some $x'$
\item Remove edge ($x$, $y$) if $x$ is dominated by some $y'$
\end{itemize}
For example, node $(W,1,1)$ in Figure~\ref{fig:example-compilation} is removed since it is never on a simple path to the end node, because the path must always go through node $(W,-,4)$. That is, node $(W,-,4)$ post-dominates node $(W,1,1)$ and both are shadows of topology location $W$.
Similarly, the edge from $(C,3,2)$ to $(D,-,2)$ is removed since node $(C,-,2)$ post-dominates $(D,-,2)$ and is a shadow of the same topology location $C$.

We repeatedly apply the minimizations above until no further minimization is possible.
In the example from Figure~\ref{fig:example-compilation}, colored nodes and dashed edges show edges and nodes removed after minimization.

@@

For example, recall the datacenter example from \S\ref{sec:motivation}, with the policy
\CD{PG1} \Path \text{ }\End(\CD{A}), where \CD{PG1} falls under the \CD{PG} aggregate. Figure~\ref{fig:aggregation-safety} shows the simplified PGIR. Since the compiler knows aggregation will occur at $X$, and it knows that the \CD{PG1} prefix will originate at $A$, we can compute the number of failures it would take to disconnect $A$ from $X$. We could remove the $A$--$D$--$X$ path first. We would then need to remove any other $A$--$D$ or $D$--$X$ links from the PGIR (in this case none). Next, we could remove the links along the $A$--$C$--$X$ path, repeating the process. Because $A$ is then disconnected from $X$, the compiler knows that 2 is a lower bound on the number of failures required to potentially introduce an aggregation black hole for prefix \CD{PG1}. This process is repeated for other aggregation locations (e.g., $Y$).

@@

Since the compiler can control community tagging only for routers under the control of the AS being programmed, it cannot match on communities for external ASes. Instead, it translates matches from external ASes into a BGP regular expression filter. For example, node $D$ in Figure~\ref{fig:example-compilation} would match the single hop external paths $X$ or $Y$. In general, if routes are allowed from beyond $X$ or $Y$, these will also be captured in the BGP regular expression filters. The unknown AS topology is modeled as a special node in the PGIR that generates a filter to match any sequence of ASes.

@@

After configuration generation, the compiler further processes the ABGP policy, removing community tags when possible, combining filters, removing dead filters, and so on. In the compilation example all community tags can be removed, since there is never any ambiguity based on the router importing the route and the neighbor the route is being imported from. Similarly, after removing the communities, the two rules at router $B$ are merged into a single rule.

@@

Our \sysname compiler is implemented in roughly 6700 lines of F\# code. It includes command-line flags for enabling or disabling the use of the BGP MED attribute, AS path prepending, the no-export community, as well as for ensuring at least k-failure safety for aggregate prefixes. Since each prefix has a separate routing policy, we compile each routing policy in parallel. Currently, \sysname supports generating Quagga router configurations out of the box. Users can add new vendor-specific adapters to translate from ABGP to other router configuration languages, or incorporate the compiler into an existing template-based system, e.g., by mixing the \sysname-generated BGP configuration with other, non-BGP configuration elements.

@@

\section*{Acknowledgments}

\ryan{TODO: add acknowledgements}
We thank R. Aditya, George Chen, and Lihua Yuan for feedback on this work and the SIGCOMM reviewers for comments on the paper.
